---
title: "GI Transit Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    theme: cosmo
    runtime: shiny
    navbar:
      - { title: "Overview", href: "#overview" }
      - { title: "Transit Time by Gender & BMI", href: "#by-factors" }
      - { title: "Transit Time vs Body Fat %", href: "#scatter-fat" }
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(ggplot2)
library(crosstalk)
library(readxl)
library(dplyr)
library(Hmisc)

# Load data
gidxa <- read_excel("clean_data/71824Parquette.xlsx")

# Data prep
gidxa$get_min <- as.numeric(gidxa$get_min)
gidxa$ctt_min_tot <- as.numeric(gidxa$ctt_min_tot)

gidxa$get_categories <- cut(gidxa$get_min, breaks = c(-Inf, 60, 240, Inf),
                            labels = c(">1 hour", "1-4 hours", "<4 hours"))
gidxa$ctt_categories <- cut(gidxa$ctt_min_tot, breaks = c(-Inf, 720, 3560, Inf),
                            labels = c(">12 hours", "12-59 hours", "<59 hours"))
gidxa$wgtt_categories <- cut(gidxa$wgtt_tot_min, breaks = c(-Inf, 600, 4320, Inf),
                             labels = c("Rapid", "Normal", "Delayed"))

gidxa$gender <- factor(gidxa$gender, levels = c("F", "M"), labels = c("Female", "Male"))

gidxa$bmi_categories <- cut(gidxa$bmi, breaks = c(-Inf, 30, 35, Inf),
                            labels = c("Overweight", "Class 1 Obesity", "Class 2 Obesity"))

# Label variables
label(gidxa$age) <- "Age"
label(gidxa$gender) <- "Gender"
label(gidxa$race) <- "Race"
label(gidxa$avg_wc) <- "Waist Circumference"
label(gidxa$bmi) <- "Body Mass Index"
label(gidxa$bl_dxa_mass) <- "Body Mass"
label(gidxa$bl_dxa_vatmass) <- "Visceral Adipose Tissue Mass"
label(gidxa$bl_dxa_vatvol) <- "Visceral Adipose Tissue Volume"
label(gidxa$MVPA_wk) <- "Moderate-Vigorous Physical Activity"
label(gidxa$get_categories) <- "Gastric Emptying Time"
label(gidxa$ctt_categories) <- "Colonic Transit Time"
label(gidxa$wgtt_categories) <- "Whole Gut Transit Time"
label(gidxa$dxa_fat_p) <- "Total Fat Percentage"

# Attach units
units(gidxa$age) <- "years"
units(gidxa$avg_wc) <- "cm"
units(gidxa$bl_dxa_mass) <- "lbs"
units(gidxa$bl_dxa_vatmass) <- "lbs"
units(gidxa$bl_dxa_vatvol) <- "in^3"
units(gidxa$MVPA_wk) <- "MET-hrs/wk"
units(gidxa$dxa_fat_p) <- "%"

# Crosstalk shared data for filtering
shared_df <- SharedData$new(gidxa, group = "GI_data")
```

## Overview {#overview}
### GI Transit and Body Composition Study

This dashboard explores how **gastrointestinal (GI) transit times** relate to **body composition** and **gender/BMI differences** in your dataset.

Use the top tabs to navigate:
- Compare **GI time by gender and BMI**
- Explore the **relationship to body fat %**

---

## Transit Time by Gender & BMI {#by-factors}

### Filters

```{r}
fluidRow(
  column(6, filter_select("select_gender", "Filter by Gender", shared_df, ~gender)),
  column(6, filter_select("select_bmi", "Filter by BMI Category", shared_df, ~bmi_categories))
)
```

### Gastric Emptying Time by Gender

```{r}
renderPlotly({
  plot_ly(gidxa, x = ~gender, y = ~get_min, type = 'box', 
          boxpoints = "outliers", jitter = 0.3, 
          pointpos = -1.8, marker = list(color = 'rgba(7, 164, 181, 0.5)')) %>%
    add_trace(type = 'scatter', mode = 'markers',
              x = ~gender, y = ~get_min,
              marker = list(size = 8, opacity = 0.7, 
                            color = ~ifelse(gender == "Female", '#E377C2', '#3E07A2')),
              hoverinfo = "text",
              text = ~paste("Gender:", gender, "<br>Transit Time:", get_min)) %>%
    layout(
      title = "Gastric Emptying Time by Gender",
      xaxis = list(title = "Gender"),
      yaxis = list(title = "Transit Time (minutes)"),
      boxmode = "group",
      hovermode = "closest"
    )
})
```

### Gastric Emptying Time by BMI Category

```{r}
renderPlotly({
  plot_ly(gidxa, x = ~bmi_categories, y = ~get_min, type = 'box', 
          boxpoints = "outliers", jitter = 0.3, 
          pointpos = -1.8, marker = list(color = 'rgba(255, 127, 14, 0.5)')) %>%
    layout(
      title = "Gastric Emptying Time by BMI Category",
      xaxis = list(title = "BMI Category"),
      yaxis = list(title = "Transit Time (minutes)"),
      hovermode = "closest"
    )
})
```

---

## Transit Time vs Body Fat % {#scatter-fat}

### GI Transit Time vs Fat %

This scatterplot shows **gastric emptying time (in minutes)** vs **total body fat percentage**. Points are colored by **gender** and faceted by **BMI category**.

```{r}
filtered_data <- reactive({
  shared_df$data(withSelection = TRUE)
})

renderPlotly({
  df <- filtered_data()

  p <- ggplot(df, aes(x = get_min, y = dxa_fat_p, color = gender)) +
    geom_point(alpha = 0.7, size = 3) +
    facet_wrap(~bmi_categories) +
    labs(title = "GI Transit Time vs Body Fat Percentage",
         x = "GI Transit Time (Minutes)",
         y = "Body Fat Percentage (%)",
         color = "Gender") +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top")

  ggplotly(p, tooltip = c("x", "y", "color")) %>%
    config(displayModeBar = FALSE) %>%
    layout(
      autosize = TRUE,
      margin = list(l = 50, r = 50, b = 100, t = 50)
    )
})
```
